---
title: "Data Reshaping Cheatsheet"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(nycflights13))
suppressPackageStartupMessages(library(dsr))
```

===

This is a cheat sheet for __Data Reshaping__ mainly for myself, but hopefully could be the reference sheet for new R programmers.
For simple reshaping, the following functions from **tidyr** are the following:

tidyr Function | Definition
----------------|------------------
   `gather`      | takes multiple columns and collapses into key values pairs
   `spread`       |  Spread a key-value pair across multiple columns
   

## Demo for these two main functions in `tidyr`

This demo is first presented using database **nycflights13**, which is the flights departing NYC in 2013. Several example of usage are shown below:

There are several tibbles in package **nycflights13**: `flights`,`weather`,`planes`,`airports`,`airlines`.
For further information of this dataset can refer to: [**nycflights13**](https://cran.r-project.org/web/packages/nycflights13/nycflights13.pdf) 

# 1. gather() function

First we look into the tibble `airports` and to express the function gather(). In this tibbles, there are 2 variables lattitude and longtitude. In this example, we will gather these two position information into one column "GPS coordinate". This could allows researcher to compare the location of the airport latitudinally or longitudinally.
```{r}
head(airports)

gather(airports,key = "GPS_coordinate", value="coordinate_in_degree", lon, lat) %>% 
  group_by(GPS_coordinate)
 
summary(dsr)

```







